FictionEro - Data Cleaning

### 2024-02-04 analysis/1_cleaning.qmd
preprocessing

df <- read.csv("../data/rawdata_participants.csv")
dftask <- read.csv("../data/rawdata_task.csv") |> 
  full_join(
    df[c("Participant", "Sex", "SexualOrientation")],
    by = join_by(Participant)
    )
```

@@

# Consecutive count of participants per day (as area)
df |>
  mutate(Date = as.Date(Date, format = "%d/%m/%Y")) |> 
  group_by(Date, Experimenter) |> 
  summarize(N = n()) |> 
  ungroup() |>
  # https://bocoup.com/blog/padding-time-series-with-r
  complete(Date, Experimenter, fill = list(N = 0)) |> 
  group_by(Experimenter) |>
  mutate(N = cumsum(N)) |>
  ggplot(aes(x = Date, y = N)) +
  geom_area(aes(fill=Experimenter)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    title = "Recruitment History",
    x = "Date",
    y = "Total Number of Participants"
  ) +
  see::theme_modern()
```

@@

```{r}
data.frame(Source = df$Experimenter,
           Comments = trimws(df$Feedback_Comments)) |> 
  filter(!Comments %in% c("", "No", "Nope", "nope", "None", "na")) |> 
  gt::gt() |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE) 
```

@@

```{r}
df |>
  ggplot(aes(x=Mobile, fill=Mobile)) +
  geom_bar() +
  theme_modern()
```

@@

@@

@@

dftask |> 
  mutate(Participant = fct_reorder(Participant, RT1)) |> 
  filter(RT1 < 100) |> 
  ggplot(aes(x=RT1, fill=Participant)) +
  geom_histogram(bins=60) +
  coord_cartesian(xlim = c(0, 15)) +
  scale_fill_metro_d(guide="none") +
  theme_modern() +
  theme(axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_blank()) +
  labs(title = "Distribution of Response Time for each Participant", x="Response time (for 3 scales) per item (s)") +
  facet_wrap(~Participant, ncol=5, scales="free_y")
```

@@

dftask |> 
  filter(Relevance %in% c("Relevant", "Non-erotic")) |> 
  group_by(Participant, Type) |> 
  summarise(Arousal = mean(Arousal), 
            Valence = mean(Valence),
            Enticement = mean(Enticement),
            .groups = "drop") |>
  pivot_wider(names_from = Type, values_from = all_of(c("Arousal", "Valence", "Enticement"))) |>
  transmute(Participant = Participant,
            Arousal = Arousal_Erotic - `Arousal_Non-erotic`,
            Valence = Valence_Erotic - `Valence_Non-erotic`,
            Enticement = Enticement_Erotic - `Enticement_Non-erotic`) |>
  filter(!is.na(Arousal)) |> 
  mutate(Participant = fct_reorder(Participant, Arousal)) |> 
  pivot_longer(-Participant) |> 
  mutate(Group = ifelse(Participant %in% outliers, "Outlier", "ok")) |> 
  ggplot(aes(x=value, y=Participant, fill=value)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient2(low = "#3F51B5", mid = "#FF9800", high = "#4CAF50", midpoint = 0) +
  # scale_fill_manual(values = c("Outlier" = "red", ok = "black"), guide="none") +
  theme_bw() +
  labs(title = "Difference between Erotic and Neutral", x="Erotic - Neutral") +
  facet_wrap(~name, ncol=3, scales="free_x")
```

@@

@@

@@

### Country of Origin

@@

@@

### English and AI-Expertise

@@

@@

>>>>

df <- read.csv("../data/rawdata_participants.csv") |> 
  mutate(across(everything(), ~ifelse(.x == "", NA, .x))) 
dftask <- read.csv("../data/rawdata_task.csv") |> 
  full_join(
    df[c("Participant", "Sex", "SexualOrientation")],
    by = join_by(Participant)
    )
```

@@

# Consecutive count of participants per day (as area)
df |>
  mutate(Date = as.Date(Date, format = "%d/%m/%Y")) |> 
  group_by(Date, Language, Experimenter) |> 
  summarize(N = n()) |> 
  ungroup() |>
  # https://bocoup.com/blog/padding-time-series-with-r
  complete(Date, Language, Experimenter, fill = list(N = 0)) |> 
  group_by(Language, Experimenter) |>
  mutate(N = cumsum(N)) |>
  ggplot(aes(x = Date, y = N)) +
  geom_area(aes(fill=Experimenter)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    title = "Recruitment History",
    x = "Date",
    y = "Total Number of Participants"
  ) +
  facet_wrap(~Language, nrow=2) +
  see::theme_modern()

```

@@

```{r}
data.frame(Language = df$Language,
           Source = df$Experimenter,
           Comments = trimws(df$Feedback_Comments)) |> 
  filter(!Comments %in% c("", "No", "Nope", "nope", "None", "na")) |> 
  arrange(Language, Source) |>
  gt::gt() |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE) 
```

@@

```{r}
df |>
  ggplot(aes(x=Mobile, fill=Mobile)) +
  geom_bar() +
  geom_hline(yintercept=nrow(df)/2, linetype="dashed") +
  theme_modern()
```

@@

The experiment's median duration is `r insight::format_value(median(df$Experiment_Duration))` min (`r print(bayestestR::hdi(df$Experiment_Duration, ci=0.5))`).

@@

```{r}
plot_hist <- function(dat) {
  dat |> 
    mutate(Participant = fct_reorder(Participant, RT1)) |> 
    filter(RT1 < 100) |> 
    ggplot(aes(x=RT1, fill=Participant)) +
    geom_histogram(bins=140) +
    coord_cartesian(xlim = c(0, 15)) +
    scale_fill_metro_d(guide="none") +
    theme_modern() +
    theme(axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_blank(),
          axis.line.y = element_blank()) +
    labs(title = "Distribution of Response Time for each Participant", x="Response time (for 3 scales) per stimuli (s)") +
    facet_wrap(~Participant, ncol=5, scales="free_y")
}
```

::: {.panel-tabset}

#### 1-60

@@

plot_hist(dftask[dftask$Participant %in% df$Participant[1:60], ])
```

#### 61-120

plot_hist(dftask[dftask$Participant %in% df$Participant[61:120], ])
```

:::

@@

dftask |> 
  filter(Relevance %in% c("Relevant", "Non-erotic")) |> 
  group_by(Participant, Type) |> 
  summarise(Arousal = mean(Arousal), 
            Valence = mean(Valence),
            Enticement = mean(Enticement),
            .groups = "drop") |>
  pivot_wider(names_from = Type, values_from = all_of(c("Arousal", "Valence", "Enticement"))) |>
  transmute(Participant = Participant,
            Arousal = Arousal_Erotic - `Arousal_Non-erotic`,
            Valence = Valence_Erotic - `Valence_Non-erotic`,
            Enticement = Enticement_Erotic - `Enticement_Non-erotic`) |>
  filter(!is.na(Arousal)) |> 
  mutate(Participant = fct_reorder(Participant, Arousal)) |> 
  pivot_longer(-Participant) |> 
  mutate(Group = ifelse(Participant %in% outliers, "Outlier", "ok")) |> 
  ggplot(aes(x=value, y=Participant, fill=value)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient2(low = "#3F51B5", mid = "#FF9800", high = "#4CAF50", midpoint = 0) +
  # scale_fill_manual(values = c("Outlier" = "red", ok = "black"), guide="none") +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  labs(title = "Difference between Erotic and Neutral", x="Erotic - Neutral") +
  facet_wrap(~name, ncol=3, scales="free_x")
```

@@

#### Sample

@@

#### Task Behaviour

```{r}
#| fig-width: 13
#| fig-height: 13
#| warning: false

show_distribution <- function(dftask, target="Arousal") {
  dftask |> 
    filter(SexualOrientation %in% c("Heterosexual", "Bisexual", "Homosexual")) |>
    bayestestR::estimate_density(select=target, 
                                 at=c("Relevance", "Category", "Sex", "SexualOrientation"), 
                                 method="KernSmooth") |>
    ggplot(aes(x = x, y = y)) +
    geom_line(aes(color = Relevance, linetype = Category)) +
    facet_grid(SexualOrientation~Sex)  +
    scale_color_manual(values = c("Relevant" = "red", "Non-erotic" = "blue", "Irrelevant"="orange")) +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_minimal()  +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          plot.title = element_text(face="bold")) +
    labs(title = target) 
}

(show_distribution(dftask, "Arousal") | show_distribution(dftask, "Valence")) /
  (show_distribution(dftask, "Enticement") | show_distribution(dftask, "Realness")) +
  patchwork::plot_layout(guides = "collect") +
  patchwork::plot_annotation(title = "Distribution of Appraisals depending on the Sexual Profile",
                             theme = theme(plot.title = element_text(hjust = 0.5, face="bold"))) 
```

We kept only heterosexual participants (`r insight::format_percent(sum(df$SexualOrientation == "Heterosexual") / nrow(df))`).

df <- filter(df, SexualOrientation == "Heterosexual")
dftask <- filter(dftask, Participant %in% df$Participant)
```

@@

::: {.panel-tabset}

### Origin

@@

```{r p_ethnicity}
ggwaffle::waffle_iron(df, ggwaffle::aes_d(group = Ethnicity), rows=10) |> 
  ggplot(aes(x, y, fill = group)) + 
  ggwaffle::geom_waffle() + 
  coord_equal() + 
  scale_fill_flat_d() + 
  ggwaffle::theme_waffle() +
  labs(title = "Self-reported Ethnicity", subtitle = "Each square represents a participant", fill="")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2)),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

@@

### Expertise

@@

:::

@@

## Beliefs about Artificial Information Technology (BAIT)

This section pertains to the validation of the BAIT scale measuring beliefs and expectations about artificial creations.

::: {.panel-tabset}

### Structure

#### Exploratory Factor Analysis

```{r}
bait <- df |> 
  select(starts_with("BAIT_"), -BAIT_Duration)

cor <- correlation::correlation(bait, redundant = TRUE) |> 
  correlation::cor_sort() |> 
  correlation::cor_lower()

clean_labels <- function(x) {
  x <- str_remove_all(x, "BAIT_") |> 
    str_replace_all("_", " - ")
}

cor |> 
  mutate(val = paste0(insight::format_value(r), format_p(p, stars_only=TRUE))) |>
  mutate(Parameter2 = fct_rev(Parameter2)) |>
  mutate(Parameter1 = fct_relabel(Parameter1, clean_labels),
         Parameter2 = fct_relabel(Parameter2, clean_labels)) |> 
  ggplot(aes(x=Parameter1, y=Parameter2)) +
  geom_tile(aes(fill = r), color = "white") +
  geom_text(aes(label = val), size = 3) +
  labs(title = "Correlation Matrix",
       subtitle = "Beliefs about Artificial Information Technology (BAIT)") +
  scale_fill_gradient2(
    low = "#2196F3",
    mid = "white",
    high = "#F44336",
    breaks = c(-1, 0, 1),
    guide = guide_colourbar(ticks=FALSE),
    midpoint = 0,
    na.value = "grey85",
    limit = c(-1, 1))  + 
  theme_minimal() +
  theme(legend.title = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

# n <- parameters::n_factors(bait, type="FA")  
# plot(n)

efa <- parameters::factor_analysis(bait, cor=cor(bait), n=2, rotation = "oblimin", sort=TRUE, scores="tenBerge", fm="ml")
plot(efa)
display(efa)
```

#### Confirmatory Factor Analysis

```{r}
m1 <- lavaan::cfa(efa_to_cfa(efa, threshold="max"), data=df)
m2 <- lavaan::cfa(
  "G =~ BAIT_5_ImitatingReality + BAIT_6_EnvironmentReal + BAIT_4_VideosIssues + BAIT_8_TextIssues + BAIT_3_VideosRealistic + BAIT_1_ImagesRealistic + BAIT_2_ImagesIssues + BAIT_7_TextRealistic", 
  data=df)

# bayestestR::bayesfactor_models(m1, m2)
lavaan::anova(m1, m2)
```
```{r}
display(parameters::parameters(m1))
```

### Scores

# Merge with data
scores <- lavaan::predict(m1) |> 
  as.data.frame() |> 
  data_rename(c("ML2"), c("BAIT_SEM")) |> 
  mutate(Participant = df$Participant) |>
  mutate(BAIT = rowMeans(select(bait, -contains("Videos")), na.rm = TRUE))

df <- full_join(df, scores[c("Participant", "BAIT_SEM", "BAIT")], by="Participant")
```

We computed two type of general scores for the BAIT scale, an **empirical** score based on the average of observed data (of the most loading items) and a **model-based** score as predicted by the structural model. The first one gives equal weight to all items (and keeps the same [0-1] range), while the second one is based on the factor loadings and the covariance structure.

```{r}
correlation::cor_test(scores, "BAIT_SEM", "BAIT") |> 
  plot() +
  ggside::geom_xsidedensity(aes(x=BAIT_SEM), color="grey", linewidth=1) +
  ggside::geom_ysidedensity(aes(y=BAIT), color="grey", linewidth=1) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .1) 
```

### Validity

#### Corrrelation with GAAIS

```{r}
table <- correlation::correlation(scores, 
                         select(df, starts_with("GAAIS")),
                         bayesian=TRUE) |> 
  summary()

format(table) |> 
  datawizard::data_rename("Parameter", "Variables") |> 
  gt::gt() |> 
  gt::cols_align(align="center") |> 
  gt::tab_options(column_labels.font.weight="bold")
```

#### AI-Expertise

```{r}
df |> 
  ggplot(aes(x=as.factor(AI_Knowledge), y=BAIT)) +
  geom_boxplot()

# m <- betareg::betareg(BAIT ~ AI_Knowledge, data=df)
m <- lm(BAIT_SEM ~ AI_Knowledge, data=df)
# m <- brms::brm(BAIT ~ mo(AI_Knowledge), data=dfsub, algorithm = "meanfield")
# m <- brms::brm(BAIT ~ AI_Knowledge, data=dfsub, algorithm = "meanfield")
display(parameters::parameters(m))

marginaleffects::predictions(m, by=c("AI_Knowledge"), newdata = "marginalmeans") |> 
  as.data.frame() |> 
  ggplot(aes(x=AI_Knowledge, y=estimate)) +
  geom_jitter2(data=df, aes(y=BAIT_SEM), alpha=0.2, width=0.1) +
  geom_line(aes(group=1), position = position_dodge(width=0.2)) +
  geom_pointrange(aes(ymin = conf.low, ymax=conf.high), position = position_dodge(width=0.2))
```

#### Gender and Age

```{r}
# m <- betareg::betareg(BAIT ~ Sex / Age, data=df, na.action=na.omit)
m <- lm(BAIT_SEM ~ Sex / Age, data=df)
display(parameters::parameters(m))
```

#### Belief in the Instructions

```{r}
glm(Feedback_LabelsIncorrect ~ BAIT_SEM * AI_Knowledge, 
    data=mutate(df, Feedback_LabelsIncorrect = ifelse(Feedback_LabelsIncorrect=="True", 1, 0)), 
    family="binomial") |> 
  parameters::parameters() |> 
  display(title="Predicting 'Labels are Incorrect'")
```

```{r}
glm(Feedback_LabelsReversed ~ BAIT_SEM * AI_Knowledge, 
    data=mutate(df, Feedback_LabelsReversed = ifelse(Feedback_LabelsReversed=="True", 1, 0)), 
    family="binomial") |> 
  parameters::parameters() |> 
  display(title="Predicting 'Labels are reversed'")
```

```{r}
glm(Feedback_CouldDiscriminate ~ BAIT_SEM * AI_Knowledge, 
    data=mutate(df, Feedback_CouldDiscriminate = ifelse(Feedback_CouldDiscriminate=="True", 1, 0)), 
    family="binomial") |> 
  parameters::parameters() |> 
  display(title="Predicting 'Easy to discriminate'")
```

```{r}
glm(Feedback_CouldNotDiscriminate ~ BAIT_SEM * AI_Knowledge, 
    data=mutate(df, Feedback_CouldNotDiscriminate = ifelse(Feedback_CouldNotDiscriminate=="True", 1, 0)), 
    family="binomial") |> 
  parameters::parameters() |> 
  display(title="Predicting 'Hard to discriminate'")
```

:::
